Apparent preferences for cognitive effort fade when multiple forms of effort

### 2021-10-14 Manuscript.Rmd
CABN revision

  Foraging has recently attracted high scientific interest due to its ecological validity, which is rooted in evolutionarily conserved choice behaviors [@Hayden2018; @Mobbs2018]. The prey selection foraging paradigm allows for the estimation of optimal, reward maximizing choice patterns on the basis of the opportunity cost of time [@Charnov1976; @Krebs1977]. The opportunity cost depends on the richness of the environment, which can be experimentally manipulated by balancing the amount of time it takes to harvest an offer (*handling time*) and to search for a new one (*travel time*). We can then interpret observed deviations from such optimal behavior, and examine how foraging behavior changes when the handling time is filled with cognitive or physical effort in comparison with an unfilled delay (thus disentangling effort from delay duration). In contrast to multi-alternative economic choice paradigms, which often treat choices as independent events, the sequential nature of foraging encourages individuals to consider both focal and global elements of the decision context. Extensive previous work on modeling foraging decisions provides computational tools to formalize relevant decision variables [@Stephens1986].

  Across two behavioral experiments, we tested the overarching hypothesis that preferences would vary depending on whether individuals faced a single type or multiple types of behavioral demands (i.e. cognitive effort, physical effort, and delay). Experiment 1 used a between-subject design in which each participant foraged for rewards of varying magnitudes that imposed varying durations of a single form of demand. Experiment 2 used a within-subject design in which individuals foraged for varying forms of demand in addition to varying reward magnitudes. First, we examined whether the discounting effects consistently observed in multi-alternative economic choice paradigms persisted when these options were presented serially, and whether the resulting foraging patterns depended on the demands imposed to obtain rewards. The existence of a well-defined optimal, reward-maximizing choice strategy helped us identify when demands boosted or detracted from the value of rewards (signified by tendencies to overharvest poor offers or underharvest advantageous offers, respectively). Second, we probed whether the perceived cost (or value) of demands was intrinsic or context-dependent. Varying the level of exposure to multiple types of demands across experiments allowed us to assess whether cost evaluation was dependent on the availability of alternative avenues of action. Further, including cognitive demands and physical demands let us examine whether such effects might apply differently to distinct types of effort.

  Our results provided mixed support for our pre-registered hypotheses, which are detailed in https://osf.io/2rsgm/registrations. For Experiment 1, we hypothesized that 1) decision makers would integrate reward and timing information in order to approximate optimal foraging strategies, as reflected by acceptance rates; 2) acceptance rates would be higher for trials that demanded physical effort and cognitive effort than for unfilled delay or effortless engagement (based on results from pilot data and previous findings from single-alternative choice paradigms; e.g., Olivola & Shafir, 2013. See also Eisenberger, 1992; Inzlicht et al., 2018); and 3) acceptance rates for each type of demand would remain stable over time, and participants would rarely exhibit within-trial reversals (accepting a trial but then quitting without receiving the reward). Consistent with our predictions, we found that individuals could approximate optimality by favoring higher rewards and shorter delays. We also found that people in the Cognitive group accepted more trials than optimal, whereas Physical and Wait groups under-accepted, partially matching our predictions but raising the possibility that cognitive effort might add value to outcomes (or attenuate subjective temporal costs). Finally, acceptance rates (and differences in acceptance rates across demands) were stable within each cost condition, in line with expectations. These findings motivated the design of Experiment 2, in which effort and delay trials were intermixed so that rejecting a trial of one type could potentially lead to a trial of the other type. For Experiment 2, we hypothesized that 1) acceptance rates would once again approximate optimality, and would show effects of reward magnitude similar to those seen in Experiment 1; 2) in contrast to Experiment 1, participants would accept unfilled-delay trials at a higher rate than trials with either form of effort; and 3) that choices would be consistent and stable over time. As before, decision makers approached the optimal strategy for the foraging environment, resembling the pattern observed in Experiment 1. Contrary to our second and third hypotheses, we found that acceptance rates were initially higher for cognitive effort than for delay, but this apparent preference disappeared over time as decision makers foraged for types of demands in addition to rewards.

@@

  The study was preregistered with the Open Science Framework (https://osf.io/2rsgm/registrations). The data and code for all analyses can be found in https://github.com/ctoroserey/Cost_studies. All experimental procedures were approved by the Boston University Institutional Review Board, and written consent was acquired for all participants. For Experiment 1, we recruited individuals until the planned number of 84 eligible participants was achieved (58 Female, median age = 21, range = 18 - 31; number excluded before reaching goal = 8). The sample size was determined by means of power analysis (ANOVA), using a significance level of 0.05, power of 0.8, an effect size of f = 0.45 (estimated from a pilot study), and three groups (one for each cost type). The resulting per-group sample was 20, which we increased to 21 in order to match three possible block orders. We added an extra group of 21 participants who experienced a minimally effortful condition in order to determine whether effort or mere task engagement was driving our results.

@@

  Participants were divided into four groups. Each participant in the first three groups was assigned to one of three types of costs (cognitive effort, physical grip-force effort, or delay), and a fourth group experienced a low-effort physical task that required gripping above a negligible force level. This last condition was added to test the possibility that any differences in acceptances between effort and delay were driven by task engagement alone. Each group was unaware that other cost conditions existed. Participants exerted physical effort by gripping a handheld dynamometer (Biopac Systems, United States) using their dominant hand. Gripping requirements in the physical effort condition were calibrated at 20% of maximum voluntary contraction (acquired at the beginning of the session). The low-effort condition was equivalent except that the grip-force threshold was set to 5%. Cognitive effort entailed switching among Stroop, dot motion judgment, and flanker tasks. Task switching has previously been shown to impose subjective costs (Kool et al., 2010). In the Stroop task, one of three color names was displayed on the screen (red, blue or green), with a font color that was either congruent (e.g. the word red presented in red) or incongruent (e.g. the word red presented in blue). Participants had to select the color of the font (i.e., they had to suppress their tendency to read the word). For the motion judgment task, 100 solid white dots moved on the screen. A fraction of these dots moved coherently to the left or right, while the rest moved in random directions (coherence could be either 30% or 40%, uniformly sampled). Participants had to respond with the direction of the coherent movement. In the flanker task, rows of arrowheads pointed either to the left or the right (maximum of 3 rows, 3 to 13 arrowheads per row). Participants responded with the direction of the center arrowhead, which could point in the same or opposite direction of its neighbors. The tasks were configured so that responses always involved a left or right key press (e.g. for Stroop, two colored circles were presented at each side of the screen). During the handling time, cognitive tasks and their configurations were randomly sampled, and were presented for 1 s followed by a 1 s inter-stimulus interval. Participants were asked to respond within each task’s presentation time (i.e., there was a 1 s response deadline). Before the experiment, participants trained in each cognitive task until they correctly performed six consecutive trials of each type. While in the handling time, if participants failed to maintain an above-threshold grip during the physical effort task (following the 1 s grace period) or made two mistakes during the cognitive task, they were forced to travel and missed the reward.

@@

  In Experiment 2, we tested decision makers' preferences for cognitive effort or physical effort relative to unfilled delay delay. We modified the prey-selection foraging task, so each participant experienced every type of demand in blocks that interleaved wait trials with either physical or cognitive effort (with a fixed 10 s handling time and 6 s travel time, to allow for comparisons between experiments). We separately analyzed Wait trials from the two block types, labeling the two sets of Wait trials according to the type of effort with which they were paired (“Wait-C” for cognitive, and “Wait-P” for physical). We hypothesized that 1) reward offer amounts would influence decisions in the same way as in Experiment 1; 2) interleaving effort and delay would produce higher acceptance rates for wait trials than for cognitive or physical effort trials, in contrast to the pattern observed in Experiment 1. and 3) that choice patterns would remain stable over time and consistent within trials.

@@

""

@@

  Understanding the sporadic attractiveness of effort remains elusive. In this study, we sought to identify factors underlying this phenomenon, and tested the hypothesis that the perceived cost or value of effort could depend on whether the decision environment contains opportunities to achieve the same outcome through different means. We conducted two foraging experiments that exposed individuals to a single or intermixed types of demands (cognitive effort, physical effort, or passive delay). We found a tendency toward engaging with cognitively effortful prospects when individuals faced a single form of demand, which faded over time when other costs were also experienced.

  Our findings provided partial support for our pre-registered hypotheses (https://osf.io/2rsgm/registrations). In accord with our predictions, decision makers in both experiments successfully approximated reward-maximizing strategies, as evidenced by acceptance rates near optimal levels. This is in line with previous foraging experiments, where the main deviation from optimality has been a tendency to overharvest [@Constantino2015; @Wikenheiser2013]. In contrast, our prediction that effort requirements would lead to greater acceptance rates only held for the Cognitive group, and this tendency did not reverse as predicted in Experiment 2. These hypotheses originally stemmed from prevalent proposals that incentivized work should become more appealing when presented in isolation [@Eisenberger1992; @Inzlicht2018], but that people would opt for the easiest way to achieve desired outcomes when multiple alternatives are available [@Hull1943; @Kool2010; @Olivola2013]. Instead, our novel use of foraging paradigms, in addition to our inclusion of types of effort that are often studied separately, highlighted that reward-independent task characteristics can promote or inhibit the pursuit of effortful prospects.

  The observed choice dynamics ruled out a number of candidate explanations. For instance, if the subjective costs and benefits of effort or delay were intrinsic, we would not have expected to observe the gradual convergence in acceptances when all three forms of demand were intermixed in the same environment. This line of reasoning also speaks against a direct role of boredom in our results. The aversiveness of boredom can lead people to attempt any available activity in order to avoid it [@westgate2020; @Wilson2014]. In our paradigm, such a tendency would have been expected to lead to a persistent preference for either effortful condition. Lastly, the gradual convergence of acceptance rates across cost conditions  rules out the possibility that our sample was mostly composed of demand seekers. Some individuals express a strong need for cognition, which can bias studies of mental labor [@Cacioppo1996; @Sayali2019]. Even though the prevalence of this trait in our college-level population could have driven the preferences for cognitive effort observed in both experiments, we would have expected such individuals to maintain this preference in the face of alternatives. Another unlikely candidate is fatigue. Even though participants decreased their acceptance rates for effort in both experiments, we would have expected participants in Experiment 2 to increase engagement with Wait trials, substituting Wait trials for effort trials in order to preserve their earnings. Our observations are unlikely to be due to the risk of failing to successfully complete effortful trials, as participants remained accurate throughout both experiments, and efficacy was consistent across cost conditions and experiments [@Fromer2021]. Finally, we cannot conclude that cognitive effort gained value from its rewarding outcomes, as has been proposed in the past [@Eisenberger1996]. This phenomenon would not have been expected to benefit a single type of demand differentially because cognitive effort, physical effort, and unfilled delay were rewarded at the same rate.

  Traditional foraging models, in which a reward is weighed against the earnings foregone by pursuing it [@Charnov1976; @Constantino2015], were also not sufficient to explain our results. Instead, our results were consistent with a computational model in which each type of demand distinctly biased the estimation of the opportunity cost, and such biases could be gradually recalibrated over time as decision makers gained experience with additional demands. In this formulation, passive delay and physical effort added to the opportunity cost, whereas cognitive effort reduced it. This raises the question of how these biases arise. We hypothesize that demands altered the perceived duration of the delay interval. Immersive activities can subjectively speed up the passage of time [@Csikszentmihalyi2014], and this effect appears to scale with the amount of cognitive processing required by a task [@Wearden2016]. Moreover, adjusting for time sensitivity has been shown to improve fits for delay discounting models [@Mckerchar2009]. Therefore, even though participants knew the nominal handling times, cognitive effort might have promoted higher acceptance rates by subjectively compressing the perceived durations. We speculate that this construal of elapsed time could then be adjusted with the experience of alternative demands, as participants learned how long the 10-second handling time felt in the context of various activities. This could help explain why preferences for cognitive effort were smaller for those participants in Experiment 2 who started with the physical effort block, as they had already begun to form an estimate of the handling time before they experienced their first cognitive effort trial. The cognitive effort task might have shortened subjective durations for several reasons, either because it divided the handling time into discrete, 2 s trials, or because taxing cognitive resources distracted individuals from estimating the passage of time [@Zakay1989]. These hypotheses will have to be experimentally probed in the future.

  Relying on foraging paradigms enabled us to address several gaps in the existing literature of effort-based decision making. For example, even though engaging with a demanding offer encompasses the evaluation, pursuit, and receipt of the outcome, investigations of the pursuit and receipt stages have been limited. Many studies have sought to isolate the evaluation stage by presenting individuals with offers that are not realized until the end of the experimental session, deliberately setting aside the question of how the ongoing experience of cost and outcome affects subsequent choices [@Kable2007; @Chong2016; @Massar2015; @Apps2015a]. However, discrepancies between expected and experienced value are well known [@Kahneman2006], and choosing immediate or low-effort outcomes could be due to factors other than time preference, such as a perceived risk of being unable to redeem the future rewards, or hoping to find better offers in the near future [@Fawcett2012; @Frederick2002; @Sozou1998]. We circumvented these pitfalls by providing participants with an opportunity to actively experience the demands they accepted, and giving them full knowledge of the available options. Another challenge in comparing effort and delay is that increments in the required effort are often paired with a longer execution time. We avoided this issue by imposing consistent handling times for all demands, allowing us manipulate temporal durations and demands for effort independently. Finally, the existence of well-defined optimal strategies provided a useful reference point for evaluating choice behavior. The apparent attractiveness of cognitive effort was reflected both by a higher proportion of acceptances at the expense of earnings, and by parameter estimates capturing bias in the estimation of opportunity costs.

  A potential limitation of this work is that we examined only a single combination of cognitively effortful tasks. Future work should investigate whether other ways of operationalizing cognitive and physical effort would be more aversive or more engaging and would resultingly have different effects on decision making. A potentially relevant factor is the way in which we calibrated cognitive and physical tasks. For physical demands, we adjusted the gripping threshold to people's maximum strength, whereas we trained participants in the cognitive tasks until they reached criterion. Training might have improved response automaticity and reduced error-related risk. In addition, attributes such as the efficacy to complete a task, or the diversity of its components, have been reported to promote engagement [@Bandura2010; @Eisenberger1996]. However, hand grip tasks have frequently been used to probe physical effort in previous work [@Chong2017; @Hogan2018; @Prevost2010], and task switching has been shown to impose subjective costs [@Kool2010; @Shenhav2017]. Furthermore, the observed decrease in acceptance rates over time for both effort domains is consistent with patterns seen for other forms of effortful exertion [@Treadway2009].

>>>>

  Foraging has recently attracted high scientific interest due to its ecological validity, which is rooted in evolutionarily conserved choice behaviors [@Hayden2018; @Mobbs2018]. The prey selection foraging paradigm allows for the estimation of optimal, reward maximizing choice patterns on the basis of the opportunity cost of time [@Charnov1976; @Krebs1977]. The opportunity cost depends on the richness of the environment, which can be experimentally manipulated by balancing the amount of time it takes to harvest an offer (*handling time*) and to search for a new one (*travel time*). In nature, this balance is exemplified when a squirrel deliberates between the handling time to crack a nut open versus the time it takes to travel to a new one. We can then interpret observed deviations from such optimal behavior, and examine how foraging behavior changes when the handling time is filled with cognitive or physical effort in comparison with an unfilled delay (thus disentangling effort from delay duration). In contrast to multi-alternative economic choice paradigms, which often treat choices as independent events, the sequential nature of foraging encourages individuals to consider both focal and global elements of the decision context. Extensive previous work on modeling foraging decisions provides computational tools to formalize relevant decision variables [@Stephens1986].

  Across two behavioral experiments, we tested the overarching hypothesis that preferences would vary depending on whether individuals faced a single type or multiple types of behavioral demands (i.e. cognitive effort, physical effort, and delay). Experiment 1 used a between-subject design in which each participant foraged for rewards of varying magnitudes that imposed varying durations of a single form of demand. Experiment 2 used a within-subject design in which individuals foraged for varying forms of demand in addition to varying reward magnitudes. First, we examined whether the discounting effects consistently observed in multi-alternative economic choice paradigms persisted when these options were presented serially, and whether the resulting foraging patterns depended on the demands imposed to obtain rewards. The existence of a well-defined optimal, reward-maximizing choice strategy helped us identify when demands boosted or detracted from the attractiveness of rewards (signified by tendencies to overharvest poor offers or underharvest advantageous offers, respectively). Second, we probed whether the perceived cost (or value) of demands was intrinsic or context-dependent. Varying the level of exposure to multiple types of demands across experiments allowed us to assess whether cost evaluation was dependent on the availability of alternative avenues of action. Further, including cognitive demands and physical demands let us examine whether such effects might apply differently to distinct types of effort.

  Our results provided mixed support for our pre-registered hypotheses, which are detailed in https://osf.io/2rsgm/registrations. For Experiment 1, we hypothesized that 1) decision makers would integrate reward and timing information in order to approximate optimal foraging strategies, as reflected by acceptance rates (based on previous reports of human foragers that approximate optimality, albeit with a tendency to overharvest; e.g. Constantino & Daw, 2015; Garrett & Daw, 2020); 2) acceptance rates would be higher for trials that demanded physical effort and cognitive effort than for unfilled delay or effortless engagement (based on results from pilot data and previous findings from single-alternative choice paradigms; e.g., Olivola & Shafir, 2013. See also Eisenberger, 1992; Inzlicht et al., 2018); and 3) acceptance rates for each type of demand would remain stable over time, and participants would rarely exhibit within-trial reversals (accepting a trial but then quitting without receiving the reward). This prediction stemmed from the intuition that stable and consistent choices would signal an understanding of the task, which we expected our participants to achieve, and which has been demonstrated in the past (see Hypothesis 1). Consistent with our predictions, we found that individuals could approximate optimality by favoring higher rewards and shorter delays. We also found that people in the Cognitive group accepted more trials than optimal, whereas Physical and Wait groups under-accepted, partially matching our predictions but raising the possibility that cognitive effort might add value to outcomes (or attenuate subjective temporal costs). Finally, acceptance rates (and differences in acceptance rates across demands) were stable within each cost condition, in line with expectations. These findings motivated the design of Experiment 2, in which effort and delay trials were intermixed so that rejecting a trial of one type could potentially lead to a trial of the other type. For Experiment 2, we hypothesized that 1) acceptance rates would once again approximate optimality, and would show effects of reward magnitude similar to those seen in Experiment 1; 2) in contrast to Experiment 1, participants would accept unfilled-delay trials at a higher rate than trials with either form of effort (given that people often opt for the easiest way to achieve a desired outcome; e.g. Hull, 1943; Kool et al., 2010; Olivola & Shafir, 2013); and 3) that choices would once again be consistent and stable over time. As before, decision makers approached the optimal strategy for the foraging environment, resembling the pattern observed in Experiment 1. Contrary to our second and third hypotheses, we found that acceptance rates were initially higher for cognitive effort than for delay, but this apparent preference disappeared over time as decision makers foraged for types of demands in addition to rewards.

@@

  The study was preregistered with the Open Science Framework (https://osf.io/2rsgm/registrations). The data and code for all analyses can be found in https://github.com/ctoroserey/Cost_studies. All experimental procedures were approved by the Boston University Institutional Review Board, and written consent was acquired for all participants. For Experiment 1, we recruited individuals until the planned number of 84 eligible participants was achieved (58 Female, median age = 21, range = 18 - 31; number excluded before reaching goal = 8). The sample size was determined by means of power analysis (ANOVA), using a significance level of 0.05, power of 0.8, an effect size of f = 0.45 (estimated from a pilot study), and three groups (one for each cost type). The resulting per-group sample was 20, which we increased to 21 in order to match three possible block orders. Therefore, this power analysis gave us an initial sample size goal of 63 total participants. We then added an extra group of 21 participants who experienced a minimally effortful condition in order to determine whether effort or mere task engagement was driving our results.

@@

  Participants were divided into four groups. Each participant in the first three groups was assigned to one of three types of costs (cognitive effort, physical grip-force effort, or delay), and a fourth group experienced a low-effort physical task that required gripping above a negligible force level. This last condition was added to test the possibility that any differences in acceptances between effort and delay were driven by task engagement alone. Each group was unaware that other cost conditions existed. Participants exerted physical effort by gripping a handheld dynamometer (Biopac Systems, United States) using their dominant hand. Gripping requirements in the physical effort condition were calibrated at 20% of maximum voluntary contraction (acquired at the beginning of the session). Participants could maintain any force level above this threshold to remain in the trial. The low-effort condition was equivalent except that the grip-force threshold was set to 5%. Cognitive effort entailed switching among Stroop, dot motion judgment, and flanker tasks. Task switching has previously been shown to impose subjective costs (Kool et al., 2010). In the Stroop task, one of three color names was displayed on the screen (red, blue or green), with a font color that was either congruent (e.g. the word red presented in red) or incongruent (e.g. the word red presented in blue). Participants had to select the color of the font (i.e., they had to suppress their tendency to read the word). For the motion judgment task, 100 solid white dots moved on the screen. A fraction of these dots moved coherently to the left or right, while the rest moved in random directions (coherence could be either 30% or 40%, uniformly sampled). Participants had to respond with the direction of the coherent movement. In the flanker task, rows of arrowheads pointed either to the left or the right (maximum of 3 rows, 3 to 13 arrowheads per row). Participants responded with the direction of the center arrowhead, which could point in the same or opposite direction of its neighbors. The tasks were configured so that responses always involved a left or right key press (e.g. for Stroop, two colored circles were presented at each side of the screen). During the handling time, cognitive tasks and their configurations were randomly sampled, and were presented for 1 s followed by a 1 s inter-stimulus interval. Participants were asked to respond within each task’s presentation time (i.e., there was a 1 s response deadline). Before the experiment, participants trained in each cognitive task until they correctly performed six consecutive trials of each type. While in the handling time, if participants failed to maintain an above-threshold grip during the physical effort task (following the 1 s grace period) or made two mistakes during the cognitive task, they were forced to travel and missed the reward.

@@

  In Experiment 2, we tested decision makers' preferences for cognitive effort or physical effort relative to unfilled delay. We modified the prey-selection foraging task, so each participant experienced every type of demand in blocks that interleaved wait trials with either physical or cognitive effort (with a fixed 10 s handling time and 6 s travel time, to allow for comparisons between experiments). We separately analyzed Wait trials from the two block types, labeling the two sets of Wait trials according to the type of effort with which they were paired (“Wait-C” for cognitive, and “Wait-P” for physical). We hypothesized that 1) reward offer amounts would influence decisions in the same way as in Experiment 1; 2) interleaving effort and delay would produce higher acceptance rates for wait trials than for cognitive or physical effort trials, in contrast to the pattern observed in Experiment 1. and 3) that choice patterns would remain stable over time and consistent within trials.

@@

@@

  Understanding the sporadic attractiveness of effort remains elusive. In this study, we sought to identify factors underlying this phenomenon, and tested the hypothesis that the perceived cost or value of effort could depend on whether the decision environment contains opportunities to achieve the same outcome through different means. We conducted two foraging experiments that exposed individuals to a single or intermixed types of demands (cognitive effort, physical effort, or passive delay). We found a tendency toward engaging with cognitively effortful prospects when individuals faced a single form of demand in Experiment 1, whereas an initially similar choice pattern faded over time when other costs were also experienced.

  Our findings provided partial support for our pre-registered hypotheses (https://osf.io/2rsgm/registrations). In accord with our predictions, decision makers in both experiments successfully approximated reward-maximizing strategies, as evidenced by acceptance rates near optimal levels. This is in line with previous foraging experiments, where the main deviation from optimality has been a tendency to overharvest [@Constantino2015; @Garrett2020; @Wikenheiser2013]. In contrast, our prediction that effort requirements would lead to greater acceptance rates only held for the Cognitive group, and this tendency did not reverse as predicted in Experiment 2. These hypotheses originally stemmed from prevalent proposals that incentivized work should become more appealing when presented in isolation [@Eisenberger1992; @Inzlicht2018], but that people would opt for the easiest way to achieve desired outcomes when multiple alternatives are available [@Hull1943; @Kool2010; @Olivola2013]. Instead, our novel use of foraging paradigms, in addition to our inclusion of types of effort that are often studied separately, highlighted that reward-independent task characteristics can promote or inhibit the pursuit of effortful prospects.

  The observed choice dynamics ruled out a number of candidate explanations. For instance, if the subjective costs and benefits of effort or delay were intrinsic, we would not have expected to observe the gradual convergence in acceptances when all three forms of demand were intermixed in the same environment. This line of reasoning also speaks against a direct role of boredom in our results. The aversiveness of boredom can lead people to attempt any available activity in order to avoid it [@westgate2020; @Wilson2014]. In our paradigm, such a tendency would have been expected to lead to a persistent preference for either effortful condition. Lastly, the gradual convergence of acceptance rates across cost conditions  rules out the possibility that our sample was mostly composed of demand seekers. Some individuals express a strong need for cognition, which can bias studies of mental labor [@Cacioppo1996; @Sayali2019]. Even though the prevalence of this trait in our college-level population could have driven the preferences for cognitive effort observed in both experiments, we would have expected such individuals to maintain this preference in the face of alternatives. Similarly, it is possible that people were driven to improve in the cognitive tasks, a motivation that might have boosted acceptance rates. Even though we cannot discard this possibility, the consistently high performance shown in the cognitive tasks (i.e. relatively few forced travels) speaks against this possibility. Another unlikely explanation is fatigue. Previous accounts have distinguished between short term fatigue, which operates in short term bursts that require interleaved rest, and long term fatigue, which produces a slow decay in performance that is unrecoverable [@bachinger2019; @meyniel2013; @muller2021]. On one hand, we found no sequential effects of quitting on choices, speaking against the type of periodic rest that is characteristic of short term fatigue. In terms of long term fatigue, even though participants decreased their acceptance rates for effort in both experiments (a hallmark of effortful exertion), we would have expected fatigued participants in Experiment 2 to increase engagement with Wait trials, substituting Wait trials for effort trials in order to preserve their earnings. Furthermore, our observations are unlikely to be due to the risk of failing to successfully complete effortful trials, as participants remained accurate throughout both experiments, and efficacy was consistent across cost conditions and experiments [@Fromer2021]. Finally, we cannot conclude that cognitive effort gained value from its rewarding outcomes, as has been proposed in the past [@Eisenberger1996]. This phenomenon would not have been expected to benefit a single type of demand differentially because cognitive effort, physical effort, and unfilled delay were rewarded at the same rate.

  Traditional foraging models, in which a reward is weighed against the earnings foregone by pursuing it [@Charnov1976; @Constantino2015], were also not sufficient to explain our results. Instead, our results were consistent with a computational model in which each type of demand distinctly biased the estimation of the opportunity cost, and such biases could be gradually recalibrated over time as decision makers gained experience with additional demands. In this formulation, passive delay and physical effort added to the opportunity cost, whereas cognitive effort reduced it. This raises the question of how these biases arise. We hypothesize that demands altered the perceived duration of the delay interval. Immersive activities can subjectively speed up the passage of time [@Csikszentmihalyi2014], and this effect appears to scale with the amount of cognitive processing required by a task [@Wearden2016]. Moreover, adjusting for time sensitivity has been shown to improve fits for delay discounting models [@Mckerchar2009]. Therefore, even though participants knew the nominal handling times, cognitive effort might have promoted higher acceptance rates by subjectively compressing the perceived durations. We speculate that this construal of elapsed time could then be adjusted with the experience of alternative demands, as participants learned how long the 10-second handling time felt in the context of various activities. This could help explain why preferences for cognitive effort were smaller for those participants in Experiment 2 who started with the physical effort block, as they had already begun to form an estimate of the handling time before they experienced their first cognitive effort trial. The cognitive effort task might have shortened subjective durations for several reasons, either because it divided the handling time into discrete, 2 s trials, or because taxing cognitive resources distracted individuals from estimating the passage of time [@Zakay1989]. Future iterations of the same foraging experiment could adapt the configuration of each effortful task in order to test this idea directly. For example, if participants estimated the elapsed duration of a trial by counting each sub-task within cognitive effort trials rather than by tracking time continuously, it should be possible to influence acceptance rates by manipulating the number of discrete sub-tasks performed during a fixed handling time for either cognitive or physical effort trials. Alternatively, if the biases were produced by an inability to direct cognitive resources to the estimation of elapsed time [@Zakay1989], increasing working memory load during the handling time should reduce the OC biases from our computational model. This could be done, for example, by assigning a different level of n-back to each participant.

  Our computational model deviated from classical formulations in ways that are pertinent to effort exertion. Some foraging models have incorporated caloric costs when the forager evaluates prospective rewards [@Charnov1976; @Stephens1986]. Whereas physical effort indeed demands such an energy expenditure, cognitive effort is not clearly associated with energy costs [@inzlicht2014; @Kurzban2013]. This physiological distinction could have produced the differences between cognitive and physical effort observed here. However, a number of features made caloric consumption an unlikely explanatory factor of our findings. First, the effects of behavioral costs on acceptance rates deviated from optimality in different directions (a difference in kind, not just degree). Second, unlike traditional models that focused primarily on animal behavior, the rewards in our study were non-caloric. Third, previous findings show that neither calorie counting nor glucose levels are a sufficient basis for effort cost perception, either physical or cognitive [@inzlicht2014]. 4) Wait trials did not demand energy consumption, yet had similar acceptance rates to Physical trials in both experiments.

  Relying on foraging paradigms enabled us to address several gaps in the existing literature of effort-based decision making. For example, even though engaging with a demanding offer encompasses the evaluation, pursuit, and receipt of the outcome, investigations of the pursuit and receipt stages have been limited. Many studies have sought to isolate the evaluation stage by presenting individuals with offers that are not realized until the end of the experimental session, deliberately setting aside the question of how the ongoing experience of cost and outcome affects subsequent choices [@Kable2007; @Chong2016; @Massar2015; @Apps2015a]. However, even though discounting occurs in experimental designs with online effort performance [@bonnelle2016; @leheron2019; @lockwood2017], discrepancies between expected and experienced value are well known [@Kahneman2006], which could affect choice dynamics over time. Further, choosing immediate or low-effort outcomes could be due to factors other than time preference, such as a perceived risk of being unable to redeem the future rewards, or hoping to find better offers in the near future [@Fawcett2012; @Frederick2002; @Sozou1998]. We therefore complement previous work by providing participants with an opportunity to actively experience the demands they accepted, and giving them full knowledge of the available options. Another challenge in comparing effort and delay is that increments in the required effort are often paired with a longer execution time. We avoided this issue by imposing consistent handling times for all demands, allowing us manipulate temporal durations and demands for effort independently. Finally, the existence of well-defined optimal strategies provided a useful reference point for evaluating choice behavior. The apparent attractiveness of cognitive effort was reflected both by a higher proportion of acceptances at the expense of earnings, and by parameter estimates capturing bias in the estimation of opportunity costs.

  A potential limitation of this work is that we examined only a single combination of cognitively effortful tasks, and we cannot fully ascertain the similarity of the perceived effort intensity across effort types. Previous studies have attempted to match the subjective intensity of different cost types by titrating the amount of each cost that yielded equivalent monetary discounting prior to the experiment (Massar et al., 2015). However, a similar intensity matching was only feasible in Experiment 2, since in Experiment 1 each person experienced one cost type, and we wanted to keep them unaware that any other possible actions existed. Nevertheless, the way in which we calibrated cognitive and physical tasks could have biased their level of engagement in opposite ways. For physical demands, we adjusted the gripping threshold to people's maximum strength, whereas we trained participants in the cognitive tasks until they reached criterion. Training might have improved response automaticity and reduced error-related risk. In addition, attributes such as the efficacy to complete a task, or the diversity of its components, have been reported to promote engagement [@Bandura2010; @Eisenberger1996]. However, hand grip tasks have frequently been used to probe physical effort in previous work [@Chong2017; @Hogan2018; @Prevost2010], and task switching has been shown to impose subjective costs [@Kool2010; @Shenhav2017]. Furthermore, the observed decrease in acceptance rates over time for both effort domains is consistent with patterns seen for other forms of effortful exertion [@Treadway2009] and the decay was similar for both effort conditions. Even so, future work should investigate whether other ways of operationalizing cognitive and physical effort would be more aversive or more engaging and would resultingly have different effects on decision making. For example, manipulating the continuity and engagement of the tasks (as suggested above) could further help us evaluate the possibility that the cognitive effort tasks were preferred because discrete task switching was deemed fun.
